---
title: "Maize in the eastern Caucasus"
author: "Samira Verhees" 
output:
  html_document: 
   toc: yes
date: 'Version: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning=FALSE, echo=FALSE, include=TRUE}

# packages
library(tidyverse) #data manipulation
library(DT) #neat and searchable tables
library(lingtypology) #linguistic maps

# data
maize <- read_tsv("maize.tsv")
villages <- read_tsv("~/Git/master_villages/villages.tsv")
biblib <- read_tsv("biblib.tsv")

# clean bib

maizebibs <-maize %>%
  select(BIBTEXKEY = example_source) %>%
  distinct()

bibmaize <- biblib %>%
  inner_join(maizebibs, by = "BIBTEXKEY")

bib_clean <- bibmaize %>%
  mutate(translation = case_when(TITLE_TRANSLATION != 'NA' ~ paste0(' [',TITLE_TRANSLATION,']'),
                                 TRUE ~ NA)) %>%
  mutate(author = case_when(is.na(AUTHOR) ~ EDITOR,
                            TRUE ~ AUTHOR)) %>%
  mutate(source = apply(cbind(author, ' (',YEAR,') ', TITLE, translation, '. ', ADDRESS,': ', PUBLISHER), 1, 
                        function(x) paste(x[!is.na(x)], collapse = ""))) %>%
  select(BIBTEXKEY, source)

```

## Introduction

This is a supplement to my poster for the Caucasian Linguistics conference of 11--13 Jun 2025 in Lyon (France). On this page you can view an interactive version of the map on my poster, as well as an additional map, and you can search the full dataset used for this study. You can also download the full dataset and a pdf of my poster from the repository.

## Maps

Each dot on the map represents a village. Color represents the type of term found, while the hue reflects the preciseness of the available data. The darkest colors represent data from that particular village. Translucent dots indicate that the only available data was general data about the language (e.g. Standard Chechen for most villages where Chechen is spoken). By clicking on a dot you can view the name of the village, the word and the value as represented in the legend. The control box in the top right corner allows you to toggle different layers of information: village < dialect non-toplevel 2 < dialect non-toplevel 1 < dialect toplevel. See the [East Caucasian villages dataset](https://sverhees.github.io/master_villages/maps_new.html) on dialect classifications.

Note that the preciseness of layers is relative. While village data is generally the most precise and language data is the most generalized, the (im)preciseness of layers depends on the size of the groupings and their internal diversity. Some dialect groups cover several sub-branchings and subsume a large number of idiosyncratic village idioms, which makes information about such a group very imprecise for a particular point on the map representing a village. On the other end of the spectrum is a language like Khinalugh, which is spoken in one village. In this case, the most precise and least precise layers coincide.

### Map 1. Name types

```{r, message = FALSE, warning=FALSE, echo=FALSE, include=TRUE, fig.width=10}

# split feature dataset into dialect layers and ascribe color(shade)s

# village layer
feature_villages <- maize %>%
  filter(map == "yes") %>%
  select(idiom, type, value1, value2, example) %>%
  filter(type == "village") %>%
  mutate(value1_color = case_when(value1 == 'hajj-wheat' ~ '#16a085',
                                  value1 == 'other wheat' ~ '#773e8e',
                                  value1 == 'other' ~ '#f39c12',
                                  TRUE ~ 'white'))

# dialect_nt2 layer
feature_nt2 <- maize %>%
  filter(map == "yes") %>%
  select(idiom, type, value1, value2, example) %>%
  filter(type == "dialect_nt2") %>%
  mutate(value1_color = case_when(value1 == 'hajj-wheat' ~ '#1ccaa7',
                                  value1 == 'other wheat' ~ '#aa71c1',
                                  value1 == 'other' ~ '#f5ae3d',
                                  TRUE ~ 'white'))


# dialect_nt1 layer

feature_nt1 <- maize %>%
  filter(map == "yes") %>%
  select(idiom, type, value1, value2, example) %>%
  filter(type == "dialect_nt1") %>%
  mutate(value1_color = case_when(value1 == 'hajj-wheat' ~ '#35e3c1',
                                  value1 == 'other wheat' ~ '#bf95d0',
                                  value1 == 'other' ~ '#f7c36e',
                                  TRUE ~ 'white'))

# dialect_toplevel layer

feature_toplevel <- maize %>%
  filter(map == "yes") %>%
  select(idiom, type, value1, value2, example) %>%
  filter(type == "dialect_toplevel") %>%
  mutate(value1_color = case_when(value1 == 'hajj-wheat' ~ '#a5f3e3',
                                  value1 == 'other wheat' ~ '#d5b8e0',
                                  value1 == 'other' ~ '#fad79e',
                                  TRUE ~ 'white'))

# general language layer

feature_lang <- maize %>%
  filter(map == "yes") %>%
  select(idiom, lang, type, value1, value2, example) %>%
  filter(type == "lang") %>%
  mutate(value1_color = case_when(value1 == 'hajj-wheat' ~ '#16a085',
                                  value1 == 'other wheat' ~ '#773e8e',
                                  value1 == 'other' ~ '#f39c12',
                                  TRUE ~ 'white'))

# combine layers for the map

# subset coordinate data
coord <- villages %>%
  select(village, lat, lon, lang, gltc_lang, dialect_toplevel, dialect_nt1, dialect_nt2, village_dialect)

# combine dialect layers and coordinates
coord_features <- coord %>%
  left_join(feature_villages, by = c("village_dialect" = "idiom")) %>%
  left_join(feature_nt2, by = c("dialect_nt2" = "idiom")) %>%
  mutate(value1 = coalesce(value1.x, value1.y),
         value2 = coalesce(value2.x, value2.y),
         example = coalesce(example.x, example.y),
         type = coalesce(type.x, type.y),
         value1_color = coalesce(value1_color.x, value1_color.y)) %>%
  select(village, lat, lon, lang, gltc_lang, dialect_toplevel, dialect_nt1, dialect_nt2, village_dialect,
         value1, value2, type, example, value1_color) %>%
  left_join(feature_nt1, by = c("dialect_nt1" = "idiom")) %>%
  mutate(value1 = coalesce(value1.x, value1.y),
         value2 = coalesce(value2.x, value2.y),
         example = coalesce(example.x, example.y),
         type = coalesce(type.x, type.y),
         value1_color = coalesce(value1_color.x, value1_color.y)) %>%
  select(village, lat, lon, lang, gltc_lang, dialect_toplevel, dialect_nt1, dialect_nt2, village_dialect,
         value1, value2, type, example, value1_color) %>%
  left_join(feature_toplevel, by = c("dialect_toplevel" = "idiom")) %>%
  mutate(value1 = coalesce(value1.x, value1.y),
         value2 = coalesce(value2.x, value2.y),
         example = coalesce(example.x, example.y),
         type = coalesce(type.x, type.y),
         value1_color = coalesce(value1_color.x, value1_color.y)) %>%
  select(village, lat, lon, lang, gltc_lang, dialect_toplevel, dialect_nt1, dialect_nt2, village_dialect,
         value1, value2, type, example, value1_color) %>%
  left_join(feature_lang, by = c("lang" = "lang")) %>%
  mutate(value1 = coalesce(value1.x, value1.y),
         value2 = coalesce(value2.x, value2.y),
         example = coalesce(example.x, example.y),
         type = coalesce(type.x, type.y),
         value1_color = coalesce(value1_color.x, value1_color.y)) %>%
  select(village, lat, lon, lang, gltc_lang, idiom, dialect_toplevel, dialect_nt1, dialect_nt2, village_dialect,
         value1, value2, type, example, value1_color) %>%
  mutate(value1_type = paste(value1, type, sep=" - ")) %>%
  filter(!is.na(type)) %>%
  mutate(type_name = case_when(type == "lang" ~ idiom,
                           type == "village" ~ village_dialect,
                           type == "dialect_toplevel" ~ dialect_toplevel,
                           type == "dialect_nt1" ~ dialect_nt1,
                           type == "dialect_nt2" ~ dialect_nt2)) %>%
  select(village, lat, lon, lang, gltc_lang, 
         value1, value2, type, example, value1_type, value1_color, type_name)

# make a subset for the general language layer

lang_features <- coord_features %>%
  filter(type=='lang')

coord_features2 <- coord_features %>%
  filter(type!='lang')

# make a map

map.feature(lang.gltc(lang_features$gltc_lang),
            latitude = lang_features$lat,
            longitude = lang_features$lon,
            features = lang_features$value1_type,
            legend = FALSE,
            control = lang_features$type,
            width = 5,
            color = lang_features$value1_color,
            opacity = 0.1,
            popup = (paste(lang_features$village,"<br>",lang_features$type_name, lang_features$type, "<br>","<i>", lang_features$example, "(",lang_features$value1_type,")")),
            zoom.control = TRUE,
            tile = "Esri.WorldGrayCanvas") %>%
map.feature(lang.gltc(coord_features2$gltc_lang),
            latitude = coord_features2$lat,
            longitude = coord_features2$lon,
            features = coord_features2$value1_type,
            legend.position = "bottomleft",
            #control = coord_features2$type,
            width = 5,
            color = coord_features2$value1_color,
            opacity = 1,
            popup = (paste(coord_features2$village,"<br>",coord_features2$type_name, coord_features2$type, "<br>","<i>", coord_features2$example, "(",coord_features2$value1_type,")")),
            zoom.control = TRUE,
            tile = "Esri.WorldGrayCanvas",
            pipe.data = .)


```

### Map 2. Name origin

```{r, message = FALSE, warning=FALSE, echo=FALSE, include=TRUE, fig.width=10}

# make a map

map.feature(lang.gltc(coord_features$gltc_lang),
            latitude = coord_features$lat,
            longitude = coord_features$lon,
            features = coord_features$value2,
            legend.position = "bottomleft",
            control = coord_features$value2,
            width = 5,
            color = "magma",
            opacity = 1,
            zoom.control = TRUE,
            popup = (paste(coord_features$village,"<br>",coord_features$type_name, coord_features$type, "<br>","<i>", coord_features$example, "(",coord_features$value2,")")),
            tile = "Esri.WorldGrayCanvas")



```


## Full dataset

Words that are set to map -- 'no' were not included in maps or counts. These include additional maize-related words, such as words for 'popcorn', as well as villages that are outside of the scope of the East Caucasian villages dataset and historical data. For data from the Dagloans database, the ID of the word list is presented in the example_page column.

```{r, message = FALSE, warning=FALSE, echo=FALSE, include=TRUE}

tableselection <- maize %>%
  select(id, family, group, lang, idiom, type, example, translation, value1, value2, map, example_source, example_page) %>%
  left_join(bib_clean, by = c("example_source" = "BIBTEXKEY")) %>%
  mutate(source = coalesce(source, example_source)) %>%
  #mutate(page = case_when(is.na(example_page) ~ '.',
   #                       TRUE ~ paste0(', ', example_page,'.'))) %>%
  select(id, family, group, lang, idiom, dialect_type = type, example, translation, name_type = value1, name_origin = value2, map, source, example_page)

datatable(tableselection, options = list(pageLength = 50, autoWidth = TRUE), filter = 'top')

```

## Type counts

```{r, message = FALSE, warning=FALSE, echo=FALSE, include=TRUE}

numbers <- maize %>%
  filter(map == "yes") %>%
  select(idiom, value1, value2) %>%
  group_by(value1, value2) %>%
  summarize(n=n())

datatable(numbers, options = list(pageLength = 13, dom = 'tip'), rownames = FALSE)

```
